Vagueness, Noise, and Signaling

### 2015-05-03 paper/version_03/paper.tex
Added reaction to suggestion for more pictures. Updated reference to "Evolving perceptual categories" and updated the text accordingly, addressing point v) of first reviewer.

Similarity-maximizing (short: sim-max) games are generalizations of
Lewis' games where different states are allowed to be more or less
similar to one another. While Lewis' games treated communicative
success as a matter of black and white, sim-max games allow for a
gradient notion: the more similar the receiver's interpretation is to
the actual state, the better. Signaling games with utility-relevant
similarities in the state space are fairly standard in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical context
for reasons that will become clear presently
\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.

@@

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi languages. Roughly put, a Voronoi
language is a pair of sender and receiver strategies, such that the
sender strategy partitions the state space into convex categories,
while the receiver's interpretations are the central spots in each
category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put,
all points in $X$ are connected via a straight line that lies entirely
in $X$; $X$ has no gaps or dents. For example, if \States is the unit
interval and all states are equiprobable, a Voronoi language with two
messages would have the sender use one message exclusively for all
points in the lower half of the unit interval and another for all
points in the upper half; the receiver's interpretations of messages
are the central points, .25 and .75, in the respective intervals
(see Figure~\ref{fig:Voronoi-language-example} for an illustration).

\begin{figure}
  \centering
  \missingfigure{Example of Voronoi language for unit interval}
  \caption{}
  \label{fig:Voronoi-language-example}
\end{figure}

@@

Seen in this light, sim-max games may provide a foundation to
approaches in cognitive semantics that rely on the notion of
conceptual spaces.  \citet[][70--77]{Gardenfors2000:Conceptual-Spac},
for example, has prominently argued that natural categories are convex
regions in conceptual space. If the conceptual space has a suitable
metric, convex categories can be derived from a set of prototypes. The
category corresponding to prototype $p$ is the set of points that are
more similar to $p$ than to any other. In this way,
\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorization system can be obtained: storing the prototypes lets us
recover the categories without having to store each category's
extension. However, what is left unexplained, is where the prototypes
come from, and why we would not see just any distribution of
prototypes as an equally efficient classification system. This is
where sim-max games can contribute a principled approach to deriving,
in an independent way, not only convex categories but also
prototypical exemplars belonging to them.

@@

These examples already show two interesting things. Firstly, inclusion
of imprecision can lead to seemingly well-behaved, yet vague
strategies in the sense that we are after. The sender strategy in
Figure~\ref{fig:example_stratsB} identifies clear positive and clear
negative cases for each signal, with a smooth transition
in-between. The receiver's interpretations of signals can be seen as
smoothed-out prototype regions. Secondly, (sender) strategies can
approach non-convex pure strategies under the replicator dynamic and
linger there for vast amounts of time, possibly indefinitely. We see
this in our limited-time simulations, but this also holds, for some
types of utility functions, for the limiting case. This was first
observed by Elliott Wagner in unpublished work
\citep[c.f.][]{OConnor2013:Evolving-Percep}. A full analysis of the
dynamics of sim-max games is beyond the scope of this paper, but we
will see shortly that diffusion from confusability of states clearly
prevents evolutionary paths that meander for a long time in the
vicinity of non-convex strategies.

>>>>

Similarity-maximizing (short: sim-max) games are generalizations of
Lewis' games where different states are allowed to be more or less
similar to one another. While Lewis' games treated communicative
success as a matter of black and white, sim-max games allow for a
gradient notion: the more similar the receiver's interpretation is to
the actual state, the better. Signaling games with utility-relevant
similarities in the state space are fairly standard in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical context
for reasons that will become clear presently.

@@

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi languages. Roughly put, a Voronoi
language is a pair of sender and receiver strategies, such that the
sender strategy partitions the state space into convex categories,
while the receiver's interpretations are the central spots in each
category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put,
all points in $X$ are connected via a straight line that lies entirely
in $X$; $X$ has no gaps or dents. For example, if \States is the unit
interval and all states are equiprobable, a Voronoi language with two
messages would have the sender use one message exclusively for all
points in the lower half of the unit interval and another for all
points in the upper half; the receiver's interpretations of messages
are the central points, .25 and .75, in the respective intervals.
See Figure~\ref{fig:Voronoi-language-example} for an illustration.

\begin{figure}
  \centering
  \missingfigure{Please include here something like Figure 4 in your paper with Elliot.}
  \caption{Example of Voronoi language for the unit interval.}
  \label{fig:Voronoi-language-example}
\end{figure}

@@

Seen in this light, sim-max games may provide a foundation to
approaches in cognitive semantics that rely on the notion of
conceptual spaces.  \citet[][70--77]{Gardenfors2000:Conceptual-Spac},
for example, has prominently argued that natural categories are convex
regions in conceptual space. If the conceptual space has a suitable
metric, convex categories can be derived from a set of prototypes. The
category corresponding to prototype $p$ is the set of points that are
more similar to $p$ than to any other. In this way,
\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorization system can be obtained: storing the prototypes lets us
recover the categories without having to store each category's
extension. However, what is left unexplained, is where the prototypes
come from, and why we would not see just any distribution of
prototypes as an equally efficient classification system. This is
where sim-max games can contribute a principled approach to deriving,
in an independent way, not only convex categories but also
prototypical exemplars belonging to them.
Development of these ideas, and more, were carried out by \citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC}.

@@

These examples already show two interesting things. Firstly, inclusion
of imprecision can lead to seemingly well-behaved, yet vague
strategies in the sense that we are after. The sender strategy in
Figure~\ref{fig:example_stratsB} identifies clear positive and clear
negative cases for each signal, with a smooth transition
in-between. The receiver's interpretations of signals can be seen as
smoothed-out prototype regions. Secondly, (sender) strategies can
approach non-convex pure strategies under the replicator dynamic and
linger there for vast amounts of time, possibly indefinitely. We see
this in our limited-time simulations, but this also holds, for some
types of utility functions, for the limiting case. This was first
observed by Elliott Wagner, as mentioned by \citet{OConnor2014-OCOEPC}.
A full analysis of the dynamics of sim-max games is beyond the scope of this paper, but we
will shortly see that diffusion from confusability of states clearly
prevents evolutionary paths that meander for a long time in the
vicinity of non-convex strategies.
